---
title: "glasseel_gam"
author:
  - Tessa Van De Hammen:
      email: babla@gmail.com
      institute: [wageningen]
      correspondence: true
  - Cédric Briand:
      institute: [iav]
  - Estibaliz Diaz:
      institute: [azti]
institute:
  - wageningen: blablabla
  - iav: Eaux & Vilaine, bd de Bretagne, La Roche Bernard 56130, France
  - azti: blablabla
documentclass: article
link-citations: yes
output:
  bookdown::word_document2:
    fig_caption: true
    number_sections: true
    global_numbering: true
    keep_md: yes
    reference_docx: "template.docx"
    pandoc_args:
      - '--lua-filter=scholarly-metadata.lua'
      - '--lua-filter=author-info-blocks.lua'
bibliography: glass_eel_size.bib
csl: "elsevier-harvard.csl"
lang: en
date: "2026-04-01"
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
datapath  <-"data"
if(Sys.info()[["user"]] == 'hamme0055'){
  setwd("C:/Tessa/glass_eel/R")
  datawd <- file.path(getwd(),datapath)
} else if (Sys.info()[["user"]] == 'cedric.briand'){
  setwd("C:/workspace/DIASPARA_glass_eel/R")
  datawd <- file.path(getwd(),datapath)
}
```

# Abstract {.unnumbered}

blabla

# Keywords {.unnumbered}

blabla *Anguilla*, glass eel, size

\newpage

# Material and method

## Data collection

Data have been collected at X different locations in Europe from Ireland
to the Mediterranean and XXX (Figure 1) . Glass eel have been analysed
alive, measured to the nearest mm in the Oria, Vilaine, Rhine Den
Oever... . In all estuaries except for the Vilaine and Irish estuaries,
glass eel have been checked for pigment stages and only lower stages
(\<=VIA3) [@elie1982] been selected for size analysis. For analysis of
weight, only stages (\<VIA2) have been used.\
In all sites glass eel have been blotted on filter paper before weight
measure.

\*\*The Vilaine\*\*

Samples collected in the Vilaine, either collected in the intraseasonal
or (later in the season) at the trapping ladder. In the intraseasonal
either sampling trips below the Arzal dam, tidal limit, 12 km from the
sea, or bought from the fishermen at the same location immediately
downstream from the dam. Glass eel are collected at the trapping ladder
throughout the year, they are more pigmented than those collected at the
same time in the estuary and this is an indication of a longer estuarine
residence [@briand2009a]. Over the sampling period, most samples
contained 50 glass eel but some were based on a lower number, especially
at the time glass eel are not numerous to reach the sample size
[@briand2003].

The Oria.

{Experimental fisheries for glass eel were carried out from 2005 to 2012
\citep{AranburuGlasseelrecruitment2015} and then sampling collection was
started anew in 2016 and 2017. Samples were collected during the new and
full moon during the fishing season. Sampling has been carried out at
four point in the Oria intraseasonal from the river mouth to the tidal
limit (12 km). From 2005 to 2009 the fishing season was
October-February. From 2009 sampling has only been carried out from
15$^{th}$ of November to January 31$^{st}$. All the glass eel in sample
were measured untill a minimum count of 30 was reached.}

The Nalon.

{The Nalon. Since 2011/12 to 2015/16 glass eel samples were collected in
experimental fisheries onboard fishing or scientific vessels. In 2016/17
and 2017/18 samples were collected in San Juan de la Arena fish market
from catches of the glass eel fishery. All samplings were carried out
during the "new moon" periods of each fishing season. All specimens were
weighed (g), measured (mm) and classified according to their
pigmentation.}

Ireland.

The glass eels sampled from the N. Ireland coastline sites of
Strangford, Carlingford and Lagan, were collected using 1m diameter drag
nets deployed along the banks of each estuary during the last half of a
rising Spring tide. These exploratory surveys were developed to provide
new index sites for direct comparative analysis against the data from
the long established glass eel collection station at the Bann. Samples
of 100+ specimens were collected across a range of dates from the
various sites between January to August 2003-2016 and preserved in 70%
ethanol. All lab based measurements (length (mm) and weight (g)) were
taken from these preserved specimens, however no determination of glass
eel developmental stage was made given that these surveys were directed
at presence/absence of glass eel.

```{r Process data, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,message=FALSE, warning=FALSE)
library(lubridate)
library(mgcv)
library(dplyr)
library(ggplot2)
library(visreg)


# read in data
load(file = file.path(datawd, "dat.Rdata"))

don   <- dat

###
origin    <- as.numeric(as.Date(ymd(19600101)))
don$fi_date <- as.Date(don$fi_date2, format= "%Y-%m-%d")

don$yday=yday(don$fi_date) # yday is the season

don$doy=don$yday-305 # year begins in november
don$season<-paste0(year(don$fi_date)-1,"-",year(don$fi_date)) # year-1-year
don$time  <- as.numeric(as.Date(don$fi_date, format= "%Y-%m-%d")) - origin



nyears <- length(unique(don$fi_year))


```

# make some exploratory plots with the raw data

```{r exploratory_plots, eval= FALSE, echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
knitr::opts_chunk$set(echo=FALSE)

# check for outliers
ggplot(don)+ geom_point(aes(x=fi_date, y=jitter(lengthmm)))+
  facet_wrap(~ser_nameshort, scales='free_y') +geom_smooth(aes(x=fi_date, y=jitter(lengthmm)))

ggplot(don)+ geom_point(aes(x=yday, y=jitter(lengthmm)))+
  facet_wrap(~ser_nameshort, scales='free_y') +geom_smooth(aes(x=yday, y=jitter(lengthmm)))

# make lf distribution
ggplot(don, aes(x=lengthmm)) +    
 geom_histogram(binwidth=1)+facet_wrap(~ser_nameshort, scales = "free_y")

```

Do some datacleaning. Remove american eels and dataset with very small
amounts of data.Run model from 2000? before it is just Den Oever.

```{r datacleaning, include=FALSE, eval=TRUE}
# remove East_river (american eel)
#don <- don[!(don$ser_nameshort %in% 'East_River'),]

# remove datasets with little data
don <- don[!(don$ser_nameshort %in% c('BdareG', 'FealGY', 'InagG', 'MaigG', 'CorGY', 'BoyneGY','LiffGY', 'InagG', 'FealGY')),]

# remove outliers Den Oever
don <- don[!(don$ser_nameshort %in% 'RhDOG' & don$yday <75),]

# remove data before 2000?
#don <- don[dat$fi_year > 1999,]
#


```

run and check the gam model

```{r gam_model, eval=FALSE}
# cr is standard cubic spline
# cc start and end at the same length (for seasonal)

don$ser_nameshort <- as.factor(don$ser_nameshort)

don1 <- don[, c("lengthmm", "ser_nameshort", "time", 'yday', 'fi_date')]
don1 <- don1[complete.cases(don1),]

g3.1 <- gam(lengthmm~ser_nameshort+s(yday,bs="cc", k=6) +
            s(time, bs = "cr", k =10),
               data=don1,
          family = gaussian)

summary(g3.1)
#gam.check(g3.1)

don1$res <- residuals(g3.1)
vis.gam(g3.1)
qq.gam(g3.1,rep=100,level=1,type="pearson",pch=19,cex=.2)

time1<-as.numeric(as.Date(ymd(20240101)))-origin

v2<-visreg(g3.1, xvar="time", cond= list(time = time1))

v1<-visreg(g3.1, xvar="yday", by= "ser_nameshort", cond= list(time = time1))
```

make plots gam model

```{r plots_gam_model, eval=FALSE, paged.print=FALSE}


ggplot(don1) + 
  facet_wrap(~ser_nameshort, scales='free_y')+
  geom_smooth(aes(x=yday, y=res))

ggplot(don1) + 
  geom_point(aes(x=yday, y=res, color=ser_nameshort))+facet_wrap(~ser_nameshort)+
  geom_smooth(aes(x=yday, y=res))

predata <- data.frame(
  dates = rep(seq(as.Date('2000-01-01'), as.Date('2025-12-31'), by = 'days'), length(unique(don$ser_nameshort))),
  ser_nameshort = rep(unique(don$ser_nameshort), each=length(seq(as.Date('2000-01-01'), as.Date('2025-12-31'), by = 'days'))))
predata$time <- as.numeric(as.Date(predata$dates, format= "%Y-%m-%d")) - origin
predata$fi_year <- format(as.Date(predata$dates, format="%Y-%m-%d"), "%Y")
predata$yday=yday(predata$dates)
predata$fi_date <- as.Date(predata$dates, format="%Y-%m-%d")

pred <- predict(g3.1, newdata=predata,se.fit=TRUE)
predata$pred_length <- pred$fit
predata$pred_length_lwr <- pred$fit-1.96*pred$se.fit
predata$pred_length_upr <- pred$fit+1.96*pred$se.fit


ggplot(predata)+
  #geom_point(aes(x=fi_date,y=jitter(lengthmm)),col="#043C6B", data=don1)+
  geom_line(aes(x=fi_date,y=pred_length), col='red')+
  geom_ribbon(data=predata,aes(x=fi_date,ymin=pred_length_lwr,ymax=pred_length_upr, group=fi_year),alpha=0.3,fill="#BF8330")+
  theme_minimal()+
  facet_wrap(~ser_nameshort, scales='free_y')
  #ylim(0,100)


p2  <-  ggplot(predata) +
  geom_line(aes(x=yday,y=pred_length, group=fi_year, color=fi_year)) +
  theme_minimal() +
  theme(panel.border = element_blank(),
        axis.line = element_line()) +
  xlab("yday") +
  facet_wrap(~ser_nameshort)+
  ylim(63,80)
print(p2)


#plot(g3.1, rug = FALSE, se = FALSE, n2 = 80, main = "gam with t2")



```

```{r gam-specific_ge_stage, eval=FALSE}


```

```{r gam-model_DenOever, eval=FALSE}
# First select Den Oever data ------------------

donrhdo <- don |> filter(ser_nameshort == "RhDOG")
donrhdo <- donrhdo[, c("lengthmm", "ser_nameshort", "time", 'yday', 'fi_date', "season")]
donrhdo <- donrhdo[complete.cases(donrhdo),]
donrhdo$year <- year(donrhdo$fi_date)
# less samples after 2002 and a hole in the series.
donrhdo |> 
  group_by(year,yday) |> 
  summarize(n=n()) |>
ggplot() + geom_point(aes(x= year, y =yday, size = n ))


donrhdo$season <- as.factor(donrhdo$season)
g0 <- gam(lengthmm~s(yday,bs="cr", k=6) +    
             s(time, bs = "cr", k = 20),
                data=donrhdo,
           family = gaussian)
plot(g0, residuals = FALSE, se = TRUE, select = 1, ylim =c(-5,5))
plot(g0, residuals = FALSE, se = TRUE, select = 2, ylim =c(-5,5))
AIC(g0) # 433328.9 Deviance explained =  24.3% => 27.5 % if k=20
# model happy to use more DF for the interannual trend if trend is common
# hierachical structure in the gam model - A general smooth and 
g1 <- gam(lengthmm~s(yday,bs="cr", k=6) + 
            s(yday,bs="cr", k=4, by = season) +
            s(time, bs = "cr", k =20),
               data=donrhdo,
          family = gaussian)
# AIC 431793.3 lower so better deviance explained 29.2% =>31.2
nvar <- donrhdo |> distinct(season) |> nrow() + 2
# plot quickly seasonal trend
plot(g1, residuals = TRUE, se = TRUE, select = 1)

# seasonal trend ---------------
# uncomment for a quick plot
#plot(g1, residuals = FALSE, se = TRUE, select = 1, ylim =c(-5,5))
# partial = don't show the partial ressiduals on the plot (to better see the trend)
# 
t2024<-as.numeric(as.Date(ymd(20240101)))-origin
v10 <- visreg(g1, xvar="yday", rug = TRUE, partial = FALSE, cond=list(time=t2024))

# Time -------------
# uncomment for a quick plot
# plot(g1, residuals = TRUE, se = TRUE, select = nvar)


v11 <- visreg(g1, xvar="time")
# get back to dates
v11$fit$time=as.Date(v11$fit$time, origin=origin)
v11$res$time=as.Date(v11$res$time, origin=origin)

plot(v11,gg=TRUE,
        points=list(cex=1.5, pch=1),
        line=list(col="darkblue"),
        fill=list(col="darkblue",fill="deepskyblue",alpha=0.3))+ 
        scale_x_date()+ theme_bw()+xlab("Date") + ylab("Size (cm)")

# all terms (to see the trend)

# plot(g1,ylim =c(-5,5))

# this one is a bit too much... 
v12 <- visreg(g1, xvar="yday", by ="season", rug = TRUE, partial = FALSE, cond=list(time=t2024))
qq.gam(g1,rep=100,level=1,type="pearson",pch=19,cex=.2)
```

```{r gam-model_east_river, eval=FALSE}
# First select east river data ------------------
# there is data from 1996 to 2019, but not later
unique(don$ser_nameshort)
donea <- don |> filter(ser_nameshort == "East_River")
donea <- donea[, c("lengthmm", "ser_nameshort", "time", 'yday', 'fi_date', "season")]
donea <- donea[complete.cases(donea),]

# in 2010 there are samples later in the season
# from 2016 there are samples earlier in the year

donea |> 
  group_by(fi_date,yday) |> 
  summarize(n=n()) |>
ggplot() + geom_point(aes(x= fi_date, y =yday, size = n ))+
  theme(axis.text.x = element_text(angle = 45, hjust=1))

donea$season <- as.factor(donea$season)

# cr is standard cubic spline
# cc start and end at the same length (for seasonal)
g10 <- gam(lengthmm~s(yday,bs="cr", k=6) +    
            s(time, bs = "cr", k =10),
               data=donea,
          family = gaussian)
AIC(g10) # 206709.9 Deviance explained = 11.3%
summary(g10)

# hierachical structure in the gam model - A general smooth and 
g11 <- gam(lengthmm~s(yday,bs="cr", k=6) + 
            s(yday,bs="cr", k=6, by = season) +
            s(time, bs = "cr", k =10),
               data=donea,
          family = gaussian) 
AIC(g11) # 203970.7 Deviance explained = 17.8%
summary(g11)

#nvar <- donea |> distinct(season) |> nrow() + 2
# plot quickly seasonal trend
plot(g10, residuals = TRUE, se = TRUE, select = 1)
plot(g10, residuals = FALSE, se = TRUE, select = 1)
plot(g10, residuals = FALSE, se = TRUE, select = 2)

plot(g11, residuals = TRUE, se = TRUE, select = 1, ylim=c(-300,1000))

plot(g11, residuals = FALSE, se = TRUE, select=1,ylim=c(-300,1000))

donea$res <- residuals(g11)
vis.gam(g10)
qq.gam(g10,rep=100,level=1,type="pearson",pch=19,cex=.2)

time1<-as.numeric(as.Date(ymd(20190101)))-origin

v10<-visreg(g10, xvar="yday", cond= list(time = time1))
v10<-visreg(g10, xvar="time", cond= list(time = time1))


v2$fit$time=as.Date(v2$fit$time, origin=as.Date(ymd(19970101)))
v2$res$time=as.Date(v2$res$time, origin=as.Date(ymd(19970101)))
v2<-visreg(g11,xvar="time")



```
