---
title: "Importing data call call glass eel length"
subtitle: "DIASPARA WP2 working document"
author: "Briand Cédric, Tessa Van Der Hamment
date: last-modified
date-format: "DD-MM-YYYY"
description: "Creation of metric db, version = build"
title-block-banner: "images/diaspara_bandeau.png"
title-block-banner-color: "white"
format:
 html:
  self-contained: true
  theme: styles.scss
  smooth-scroll: true
  fontcolor: black
  toc: true
  toc-location: left
  toc-title: Summary
  toc-depth: 3
execute: 
 keep-md: true
filters:
  - include-code-files
reference-location: document
bibliography: diaspara.bib
include-after-body: "footer.html"
---



```{r init}
#| echo: FALSE
#| warning: FALSE
#| message: FALSE
#| results: 'hide'

#if (!grepl("montepomi", getwd())) {
if(Sys.info()[["user"]] == 'joliviero'){
setwd("D:/workspace/DIASPARA_WP3_migdb/R")
datawd <- "D:/DIASPARA/wgbast"
} else if (Sys.info()[["user"]] == 'cedric.briand'){
setwd("C:/workspace/DIASPARA_WP3_migdb/R")
datawd <- "C:/Users/cedric.briand/OneDrive - EPTB Vilaine/Projets/DIASPARA/wgbast"
}
source("utilities/load_library.R")
load_library("tidyverse")
load_library("knitr")
load_library("kableExtra")
load_library("icesVocab")
load_library("readxl")
load_library("janitor")
load_library("skimr")
load_library("RPostgres")
load_library("yaml")
load_library("DBI")
load_library("ggplot2")
load_library("sf")
load_library("janitor") # clean_names
load_library("uuid")
load_library("stacomiR")
load_library("readxl")
load_library("writexl")
cred <- read_yaml("../credentials.yml")
con_diaspara <- dbConnect(Postgres(), 
                           dbname = cred$dbnamediaspara,
                           host = cred$host,
                           port = cred$port,
                           user = cred$userdiaspara,
                           password = cred$passworddiaspara)
con_diaspara_admin <- dbConnect(Postgres(), 
                           dbname = cred$dbnamediaspara,
                           host = cred$host,
                           port = cred$port,
                           user = cred$usersalmo,
                           password = cred$passwordsalmo)
con_salmoglob <- dbConnect(Postgres(), 
                           dbname = cred$dbnamesalmo,
                           host = cred$host,
                           port = cred$port,
                           user = cred$usersalmo,
                           password = cred$passwordsalmo)
con_wgeel_distant <- dbConnect(Postgres(), 
                           dbname = cred$dbnamedistant,
                           host = cred$hostdistant,
                           port = cred$port,
                           user = cred$userdistant,
                           password = cred$passworddistant)
con_wgeel_local <- dbConnect(Postgres(), 
                           dbname = "wgeel",
                           host = '127.0.0.1',
                           port = cred$port,
                           user = cred$userdistant,
                           password = cred$passwordsalmo)
if (exists("userdistant") & exists("passworddistant") & exists("hostdistant")) {
	if( !exists("pois"))  pois <- mainpass
	host <- decrypt_string(hostdistant,pois)
	user <- decrypt_string(userdistant,pois)
	password<- decrypt_string(passworddistant,pois)
} else {
  load_library('getPass')
	host <- getPass(msg="host")
	user <- getPass(msg="user")
	password <- getPass(msg="password")
}

options(list(
				stacomiR.dbname = "bd_contmig_nat_iav",
				stacomiR.host = host,
				stacomiR.port = "5432",
				stacomiR.user = user,
				stacomiR.password = password,
				stacomiR.printqueries =FALSE
		))
#printqueries mettre TRUE pour voir les requetes
stacomi(TRUE,sch="iav", datawd=datawd)


 write_xlsx(as.data.frame(list_comp_individual_metrics$new), file)
```


```{r }
#| label: TRUE
#| echo: TRUE
#| warning: FALSE
#| message: FALSE

vilG <- dbGetQuery(con_wgeel_distant, "SELECT * FROM datawg.t_series_ser WHERE ser_nameshort = 'VilG'")

#clipr::write_clip(vilG)
#LOADING TRAP DATA ------------------------------------

con <- dbConnect(Postgres(), 		
    dbname=getOption("stacomiR.dbname"), 		
    host=getOption("stacomiR.host"),
    port=getOption("stacomiR.port"), 		
    user= getOption("stacomiR.user"), 		
    password= getOption("stacomiR.password"))
tps <- dbGetQuery(con, "select * from iav.civelle_taille_poids_stade")
#colnames(tps)
tps$ser_nameshort="VilTG"
tps <- tps[,-match("ope_operateur",colnames(tps))]
tps <- tps %>% rename("fi_id_cou"="lot_identifiant",
                "fi_date"="ope_date_debut",
                "lengthmm" = taille,
                "weightg" = poids,
                "ge_stage" = stade
             )
#LOADING ESTUARINE DATA ----------------------------------------

con <- dbConnect(Postgres(), 		
    dbname="mortciv", 		
    host=getOption("stacomiR.host"),
    port=getOption("stacomiR.port"), 		
    user= getOption("stacomiR.user"), 		
    password= getOption("stacomiR.password"))

tps2 <- dbGetQuery(con, "select * from vue_ope_lot_taillepoidsstade")
tps2 <- tps2[, c("ope_identifiant","lot_identifiant","datedebutpeche","taille","poids","stade")]
tps2$ser_nameshort="VilG"
colnames(tps2) <- colnames(tps)
vil <- rbind(tps,tps2)# row bind = concatenate
save(vil,file=file.path("data","vil.Rdata"))
# !!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
# START HERE WITHOUT CONNEXION TO THE DATABASE
# load(file=str_c(datawdy,"vil.Rdata"))
#!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!

# VILAINE (data post treatment) ------------------------------

# this will change names so that they correspond to yours

vil$fi_date <- as.Date(vil$fi_date)
vil$fi_year <- lubridate::year(vil$fi_date)
vil <- vil %>% select(fi_id_cou,ser_nameshort,fi_date,fi_year,ge_stage,lengthmm,	weightg) %>% mutate(fi_comment = NA)
)
write_xlsx(x=list("new_individual_metrics"=vil), path = file.path("data,vil.xlsx")
# I paste it from there to excel
dbDisconnect(con)



# p + geom_point(size=0.5) + 
#     scale_x_date(name="date")


@
